---
title: "Weekly feeback"
author: "Simon Coburg; IMBRSea 20191199 - CCMAR a73216"
date: "11/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Thesis "Seagrass domination over seaweeds in oligotrophic coastal ecosystems"

<font size="3"> with supervisors: *[Ana Alexandre](https://www.ccmar.ualg.pt/users/aalexandre), [Carmen de los Santos](https://www.ccmar.ualg.pt/users/cbsantos) & [Rui Santos](https://www.ccmar.ualg.pt/users/rosantos), CCMAR* </font>

---

*R script - [GitHub](https://github.com/NestaKobe/Uptake_rates)*

---

PRISMA protocol step **Eligibility**

*of full-text articles assessed for eligibility:* 345

- total studies assessed at present: 345

- of full-text articles excluded with reasons: 31

- of full-text articles on waiting list: 33   

- <u> of studies included in qualitative synthesis: 88 </u>

<br>

### Exploratory data analysis
<br>
```{r loading libraries, include=FALSE}

# load libraries
packages <- c("tidyverse",      # for data science (general) - includes ggplot2
              "readxl",         # for reading xlsx files
              "devtools",
              "ggpubr")

for (i in seq_along(packages)) {
  if(!do.call(require, list(package = packages[i]))) {
    do.call(install.packages, list(pkgs = packages[i]))
    do.call(require, list(package = packages[i]))
  }
}

# set working directory
setwd("~/Documents/IMBRSea/Thesis S4/Rstudio Uptake rates")
```

```{r loading data, include=FALSE}

# load SOURCES
data.sou  <- read_excel("~/Documents/IMBRSea/Thesis S4/Database uptake rate_04.06.2021.xlsx",
                        sheet="sources",na="NA",skip=3)

str(data.sou)
names(data.sou)

# load ENVIRONMENTAL
data.env  <- read_excel("~/Documents/IMBRSea/Thesis S4/Database uptake rate_04.06.2021.xlsx",
                        sheet="environmental",na="NA",skip=3)
str(data.env)
names(data.env)

# load EXPERIMENTAL
data.exp  <- read_excel("~/Documents/IMBRSea/Thesis S4/Database uptake rate_04.06.2021.xlsx",
                        sheet="experimental",na="NA",skip=3)
str(data.exp)
names(data.exp)
```

```{r converting numeric, warning=FALSE, include=FALSE}

data.exp$Vmax   <- as.numeric(data.exp$Vmax)
data.exp$Km     <- as.numeric(data.exp$Km)
data.exp$alpha  <- as.numeric(data.exp$alpha)
data.exp$sampling_inter <- as.numeric(data.exp$sampling_inter)

```

```{r combining, waring=FALSE, include=FALSE}
# MERGE spreadsheets
data.ee <- merge(data.env,data.exp,by="id_short")
data.all <- merge(data.ee, data.sou, by="study_id")

#Filter for RANGE only
data.range <- data.all[data.all$type_incub == "Range",]
```

**Quality controls for ENVIRONMENTAL/EXPERIMENTAL**
<br>
Comparison between spreadsheets *environmental* and *experimental* whether <u> unique_id </u>
coincide.
```{r}
length(unique(data.env$id_short))
length(unique(data.exp$id_short))
```
<br>

### Dataset
<br>
Number of articles registered for *Vmax*, *Km* and *alpha* values up to date:
<br>
*PRISMA protocol step Eligibility - <u>of studies includes in qualitative synthesis</u>*
```{r sum sampling location, echo=FALSE, warning=FALSE}
genv <- unique(data.range[,c("study_id", "year")]) # create data for ggplot - 

#using the unique function with study_id and year you eliminate repeated countries within the same study
gdata <- data.frame(table(genv$year)) # this is to create a frequency table by year
names(gdata) <- c("year","n")

sum(gdata$n)

rm(gdata, genv) # delete gdata

```
<br>

**Publication year**
<br>
Peer-reviewed publications assessed by year up to date:
<br>


```{r publication year, echo=FALSE, warning=FALSE}
genv <- unique(data.range[,c("study_id", "year")]) # create data for ggplot - 

#using the unique function with study_id and year you eliminate repeated countries within the same study
gdata <- data.frame(table(genv$year)) # this is to create a frequency table by year
names(gdata) <- c("year","n")
gdata$year <- as.numeric(as.character(gdata$year))

ggplot(gdata, aes(x=year, y=n)) +
      geom_bar(stat="identity", na.rm=FALSE) +
      ggtitle("Publication year") + 
      labs(x="Year of publication", y="Number of articles") +
      scale_x_continuous(limits=c(1975, 2021), breaks=seq(1975,2021,5)) +
      scale_y_continuous(limits=c(0,8), breaks=seq(0,8,2)) +
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5)) +
      theme(axis.text.x = element_text(angle = 0))
      
rm(gdata, genv) # delete gdata
```
<br>
<br>

**Sampling locations**
<br>
Amount of peer-reveiwed articles' sampling countries up to date:
<br>

```{r sampling location, echo=FALSE, warning=FALSE}
genv <- unique(data.range[,c("study_id", "sample_country")]) # create data for ggplot - 

#using the unique function with study_id and sample_country you eliminate repeated countries within the same study
gdata <- data.frame(table(genv$sample_country)) # this is to create a frequency table for countries
names(gdata) <- c("sample_country","n")

ggplot(gdata, aes(x=reorder(sample_country,n), y=n)) +   # here I have added "reorder" so in the plot you can see countries ordered by frequency
      geom_bar(stat="identity", na.rm=FALSE) + 
      coord_flip() +
      scale_x_discrete("Country") +
      scale_y_continuous("Number of studies", limits = c(0, 25), breaks = seq(0, 25, by = 5)) +
      ##adjust scale, if numbers higher than limits set --> excluded from plot
      theme_bw()
      
rm(gdata, genv) # delete gdata
```
<br>
<br>

**Species type**
<br>
Registered values portrayed by species types *Algae* and *Seagrasses*, by *phyla* and the species *compartments*:

```{r species type, echo=FALSE}
ggplot(data.range, aes(x=species_type)) +
      geom_bar(stat="count",  width = .5) +
      geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, colour = "white") +
      ggtitle("Registered values by species type") + 
      labs(x="Species type", y="Number of registered values") +
      theme_bw()+
      theme(plot.title = element_text(hjust = 0.5))
```
<br>
<br>

```{r species phyla, echo=FALSE}
  
ggplot(data.range, aes(x=species_type, fill=species_phyla)) +
      scale_fill_manual(values = c("forestgreen", "salmon4", "red2", "springgreen4")) +
      geom_bar(stat="count", position=position_dodge(),  width = .6)+
      ggtitle("Registered values by species phyla") + 
      labs(x="Species type", y="Number of registered values") +
      scale_y_continuous(limits = c(0, 350), breaks = seq(0, 350, by =50)) + 
      #adjust scale, if numbers higher than limits set --> excluded from plot
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5), legend.position="right") +
      theme(legend.title = element_blank()) +
      theme(legend.position = "bottom")
```
<br>
<br>

```{r species compartments, echo=FALSE}

ggplot(data.range, aes(x=species_phyla, colour=species_compartm, fill=species_compartm)) +
      geom_bar(stat="count", position=position_dodge(), width= .6) +
      ggtitle("Registered values by species compartment for phyla") + 
      labs(x="", y="Number of registered values") +
      scale_y_continuous(limits = c(0, 225), breaks = seq(0, 225, by = 25)) + 
      #adjust scale, if numbers higher than limits set --> excluded from plot
      facet_grid(species_type~.) +
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5), legend.position="bottom") +
      theme(legend.title = element_blank())
```
<br>
<br>
<br>

#### Alpha values
Alpha frequency distribution of values registered up to date:

```{r alpha distribution, echo=FALSE, warning=FALSE}
ggplot(data.range, aes(x=alpha))+
      geom_histogram(binwidth=0.1, 
                     breaks = seq(0, 15, by = 0.2), 
                     colour = "black", 
                     fill = "white") +
      geom_rug() +
      geom_density(aes(y=..density..*10), colour="blue") +
      scale_x_continuous(limits=c(0, 15)) +
      scale_y_continuous(limits=c(0, 20)) +
      facet_grid(type_uptake~species_type) +
      ggtitle("Alpha frequency distribution") + 
      labs(x="Alpha [l g^-1 dw h^-1 µM^-1]", y="Frequency") +
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5))
```
<br>
<br>

*Alpha frequency distribution portrayed as overlapping densities:*

```{r alpha distribution overlapping, echo=FALSE, warning=FALSE}
#Overlapping densities
ggplot(data.range, aes(x=alpha, group=species_type, fill=species_type))+
      geom_density(alpha=0.4) +
      facet_grid(.~type_uptake) +
      ggtitle("Alpha frequency distribution for Algae vs Seagrass") + 
      labs(x="Alpha [l g^-1 dw h^-1 µM^-1]", y="Frequency") +
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5)) +
      theme(legend.title = element_blank()) +
      theme(legend.position = c(.9,.88))

```
<br>
<br>
Alpha values by nutrients:
<br>

```{r alpha distribution nutients, echo=FALSE, warning=FALSE}
ggplot(data.range, aes(x=species_type, y=alpha)) +
      geom_boxplot(outlier.shape=NA) +
      geom_jitter(width=0.3, shape=21) +
      facet_grid(type_uptake~nutrient) +
      ggtitle("Alpha values for nutrients") + 
      labs(x="Species type", y="Alpha [l g^-1 dw h^-1 µM^-1]") +
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5))
```
<br>
<br>

*Alpha frequency distribution by nutrients portrayed as overlapping densities:*

```{r alpha distribution nutrients overlapping, echo=FALSE, warning=FALSE}

#Overlapping densities
ggplot(data.range, aes(x=alpha, group=species_type, fill=species_type))+
      geom_density(alpha=0.3) +
      facet_grid(type_uptake~nutrient) +
      ggtitle("Alpha frequency distribution by nutrient for Algae vs Seagrass") + 
      labs(x="Alpha [l g^-1 dw h^-1 µM^-1]", y="Frequency") +
      scale_x_continuous(limits=c(0, 0.75), breaks=seq(0, 1, by=0.25)) +
      #scale_y_continuous(breaks = seq(0, 12, by =5)) +
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5)) +
      theme(legend.title = element_blank()) +
      theme(legend.position = "bottom") +
      theme(axis.text.x = element_text(angle = 90))
```
<br>
<br>
Alpha values by phyla and by nutrients over 60 minutes:
<br>

```{r alpha time, echo=FALSE, warning=FALSE}

#by incubation time
ggplot(data.range, aes(x=time, y=alpha, colour=nutrient)) +
      geom_point(outlier.shape=NA) +
      facet_grid(species_phyla~nutrient) +
      ggtitle("Alpha values over 60 minutes") + 
      labs(x="Time [minutes]", y="Alpha [l g^-1 dw h^-1 µM^-1]") +
      scale_x_continuous(limits=c(0, 60)) +
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5)) +
      theme(legend.position = "none")
```
<br>
<br>
**Please note:** *a file was written as* ***Outliers_alpha.csv*** *and added to this reports feedback.*

```{r alpha outliers, include=FALSE}
out <- boxplot.stats(data.range$alpha)$out #look for outliers
out_ind <- which(data.range$alpha %in% c(out))
out_ind
  
outliers <- data.range[out_ind, ]
outliers <- unique(outliers[,c("study_id", "id_short", "species", "species_type", 
                                 "species_phyla", "species_compartm", "temperature_experiment", 
                                 "lumination_experiment", "nutrient", "time", "type_uptake", 
                                 "Vmax", "Km", "alpha", "annotations_exp")])
  
write_csv(outliers, file = "Outliers_alpha.csv")
```
<br>
<br>
<br>

#### Vmax values
Vmax frequency distribution of values registered  up to date:
<br>

```{r Vmax distribution, echo=FALSE, warning=FALSE}
ggplot(data.range, aes(x=Vmax))+
      geom_histogram(binwidth=0.1, 
                     breaks = seq(0, 15, by = 0.2), 
                     colour = "black", 
                     fill = "white") +
      geom_rug() +
      geom_density(aes(y=..density..*10), colour="blue") +
      scale_x_continuous(limits=c(0, 20)) +
      scale_y_continuous(limits=c(0, 10)) +
      facet_grid(type_uptake~species_type) +
      ggtitle("Vmax frequency distribution") + 
      labs(x="Vmax [µmol g^-1 dw h^-1]", y="Frequency") +
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5))
```
<br>
<br>
*Vmax frequency distribution portrayed as overlapping densities:*

```{r Vmax distribution overlapping, echo=FALSE, warning=FALSE}
#Overlapping densities
ggplot(data.range, aes(x=Vmax, group=species_type, fill=species_type))+
      geom_density(alpha=0.4) +
      scale_x_continuous(limits=c(0,200)) +
      facet_grid(.~type_uptake) +
      ggtitle("Vmax frequency distribution for Algae vs Seagrass") + 
      labs(x="Vmax [µmol g^-1 dw h^-1]", y="Frequency") +
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5)) +
      theme(legend.title = element_blank()) +
      theme(legend.position = c(.9,.88))

```
<br>
<br>

Vmax values by nutrients:
<br>

```{r Vmax distribution nutrients, echo=FALSE, warning=FALSE}
ggplot(data.range, aes(x=species_type, y=Vmax)) +
      geom_boxplot(outlier.shape=NA) +
      geom_jitter(width=0.3, shape=21) +
      facet_grid(type_uptake~nutrient) +
      ggtitle("Vmax values for nutrients") + 
      labs(x="Species type", y="Vmax [µmol g^-1 dw h^-1]") +
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5))
```
<br>
<br>

*Vmax frequency distribution by nutrients portrayed as overlapping densities:*

```{r Vmax distribution nutrients overlapping, echo=FALSE, warning=FALSE}
#Overlapping densities
ggplot(data.range, aes(x=Vmax, group=species_type, fill=species_type))+
      geom_density(alpha=0.3) +
      facet_grid(type_uptake~nutrient) +
      ggtitle("Vmax frequency distribution by nutrient for Algae vs Seagrass") + 
      labs(x="Vmax [µmol g^-1 dw h^-1]", y="Frequency") +
      scale_x_continuous(limits=c(0, 15), breaks=seq(0, 15, by=5)) +
      #scale_y_continuous(breaks = seq(0, 1.5, by =0.5)) +
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5)) +
      theme(legend.title = element_blank()) +
      theme(legend.position = "bottom") +
      theme(axis.text.x = element_text(angle = 0))
```
<br>
<br>
Vmax values by phyla and by nutrients over 60 minutes:
<br>

```{r Vmax time, echo=FALSE, warning=FALSE}

#by incubation time
ggplot(data.range, aes(x=time, y=Vmax, colour=nutrient)) +
      geom_point(outlier.shape=NA) +
      facet_grid(species_phyla~nutrient) +
      ggtitle("Vmax values over 60 minutes") + 
      labs(x="Time [minutes]", y="Vmax [µmol g^-1 dw h^-1]") +
      scale_x_continuous(limits=c(0, 60)) +
      scale_y_continuous(limits=c(0,600)) +
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5)) +
      theme(legend.position = "none")

```
<br>
<br>
**Please note:** *a file was written as* ***Outliers_Vmax.csv*** *and added to this reports feedback.*
```{r Vmax outliers, include=FALSE}
#Outliers
out <- boxplot.stats(data.range$Vmax)$out #look for outliers
out_ind <- which(data.range$Vmax %in% c(out))
out_ind

outliers <- data.range[out_ind, ]
outliers <- unique(outliers[,c("study_id", "id_short", "species", "species_type", 
                               "species_phyla", "species_compartm", "temperature_experiment", 
                               "lumination_experiment", "nutrient", "time", "type_uptake", 
                               "Vmax", "Km", "alpha", "annotations_exp")])

write_csv(outliers, file = "Outliers_Vmax.csv")
```
<br>
<br>
<br>

**This concludes this weeks feedback file.**
<br>
<br>
If you have any suggestions or requests for other graphs and/or information, please let me know.
Best, Simon
