---
title: "Weekly feeback"
author: "Simon Coburg; IMBRSea 20191199 - CCMAR a73216"
date: "21/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Thesis "Seagrass domination over seaweeds in oligotrophic coastal ecosystems"

<font size="3"> with supervisors: *[Ana Alexandre](https://www.ccmar.ualg.pt/users/aalexandre), [Carmen de los Santos](https://www.ccmar.ualg.pt/users/cbsantos) & [Rui Santos](https://www.ccmar.ualg.pt/users/rosantos), CCMAR* </font>

---

*R script - [GitHub](https://github.com/NestaKobe/Uptake_rates)*

---

PRISMA protocol step **Eligibility**

*of full-text articles assessed for eligibility:* 345

- total studies assessed at present: 153

- of full-text articles excluded with reasons: 30

- of full-text articles on waiting list: 10   

- <u> of studies included in qualitative synthesis: 30 </u>

<br>

### Exploratory data analysis
<br>
```{r loading libraries, include=FALSE}

# load libraries
packages <- c("tidyverse",      # for data science (general) - includes ggplot2
              "readxl",         # for reading xlsx files
              "devtools",
              "ggpubr")

for (i in seq_along(packages)) {
  if(!do.call(require, list(package = packages[i]))) {
    do.call(install.packages, list(pkgs = packages[i]))
    do.call(require, list(package = packages[i]))
  }
}

# set working directory
setwd("~/Documents/IMBRSea/Thesis S4/Rstudio Uptake rates")
```

```{r loading data, include=FALSE}

# load SOURCES
data.sou  <- read_excel("~/Documents/IMBRSea/Thesis S4/Database uptake rate_21.05.2021.xlsx",
                        sheet="sources",na="NA",skip=3)

str(data.sou)
names(data.sou)

# load ENVIRONMENTAL
data.env  <- read_excel("~/Documents/IMBRSea/Thesis S4/Database uptake rate_21.05.2021.xlsx",
                        sheet="environmental",na="NA",skip=3)
str(data.env)
names(data.env)

# load EXPERIMENTAL
data.exp  <- read_excel("~/Documents/IMBRSea/Thesis S4/Database uptake rate_21.05.2021.xlsx",
                        sheet="experimental",na="NA",skip=3)
str(data.exp)
names(data.exp)
```

```{r converting numeric, warning=FALSE, include=FALSE}

data.exp$Vmax   <- as.numeric(data.exp$Vmax)
data.exp$Km     <- as.numeric(data.exp$Km)
data.exp$alpha  <- as.numeric(data.exp$alpha)
data.exp$sampling_inter <- as.numeric(data.exp$sampling_inter)

```

```{r combining, waring=FALSE, include=FALSE}
# MERGE spreadsheets
data.ee <- merge(data.env,data.exp,by="id_short")
data.all <- merge(data.ee, data.sou, by="study_id")

#Filter for RANGE only
data.range <- data.all[data.all$type_incub == "Range",]
```

**Quality controls for ENVIRONMENTAL/EXPERIMENTAL**
<br>
Comparison between spreadsheets *environmental* and *experimental* whether <u> unique_id </u>
coincide.
```{r}
length(unique(data.env$id_short))
length(unique(data.exp$id_short))
```
<br>

### Dataset
<br>
Number of articles registered for *Vmax*, *Km* and *alpha* values up to date:
<br>
*PRISMA protocol step Eligibility - <u>of studies includes in qualitative synthesis</u>*
```{r sum sampling location, echo=FALSE, warning=FALSE}
genv <- unique(data.range[,c("study_id", "year")]) # create data for ggplot - 

#using the unique function with study_id and year you eliminate repeated countries within the same study
gdata <- data.frame(table(genv$year)) # this is to create a frequency table by year
names(gdata) <- c("year","n")

sum(gdata$n)

rm(gdata, genv) # delete gdata

```
<br>

**Publication year**
<br>
Peer-reviewed publications assessed by year up to date:
<br>
*Note: I will need to work on this graph to have a continuous timeline*
<br>


```{r publication year, echo=FALSE, warning=FALSE}
genv <- unique(data.range[,c("study_id", "year")]) # create data for ggplot - 

#using the unique function with study_id and year you eliminate repeated countries within the same study
gdata <- data.frame(table(genv$year)) # this is to create a frequency table by year
names(gdata) <- c("year","n")


ggplot(gdata, aes(x=year, y=n)) +
      geom_bar(stat="identity", na.rm=FALSE) +
      ggtitle("Publication year") + 
      labs(x="Year of publication", y="Number of articles") +
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5))
      
rm(gdata, genv) # delete gdata
```
<br>
<br>

**Sampling locations**
<br>
Amount of peer-reveiwed articles' sampling countries up to date:
<br>

```{r sampling location, echo=FALSE, warning=FALSE}
genv <- unique(data.range[,c("study_id", "sample_country")]) # create data for ggplot - 

#using the unique function with study_id and sample_country you eliminate repeated countries within the same study
gdata <- data.frame(table(genv$sample_country)) # this is to create a frequency table for countries
names(gdata) <- c("sample_country","n")

ggplot(gdata, aes(x=reorder(sample_country,n), y=n)) +  # here I have added "reorder" so in the plot you can see countries ordered by frequency
      geom_bar(stat="identity", na.rm=FALSE) + 
      coord_flip() +
      scale_x_discrete("Country") +
      scale_y_continuous("Number of studies") +
      theme_bw() +
      

rm(gdata, genv) # delete gdata
```
<br>
<br>

**Species type**
<br>
Registered values portrayed by species types *Algae* and *Seagrasses*, by *phyla* and the species *compartments*:

```{r species type, echo=FALSE}
ggplot(data.range, aes(x=species_type)) +
      geom_bar(stat="count",  width = .5) +
      geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, colour = "white") +
      ggtitle("Registered values by species type") + 
      labs(x="Species type", y="Number of registered values") +
      theme_bw()+
      theme(plot.title = element_text(hjust = 0.5))
```
<br>
<br>

```{r species phyla, echo=FALSE}
  
ggplot(data.range, aes(x=species_type, fill=species_phyla)) +
      scale_fill_manual(values = c("forestgreen", "salmon4", "red2", "springgreen4")) +
      geom_bar(stat="count", position=position_dodge(),  width = .6)+
      ggtitle("Registered values by species phyla") + 
      labs(x="Species type", y="Number of registered values") +
      scale_y_continuous(limits = c(0, 120), breaks = seq(0, 120, by = 20)) +
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5), legend.position="right") +
      theme(legend.title = element_blank()) +
      theme(legend.position = c(.85,.8))
```
<br>
<br>

```{r species compartments, echo=FALSE}

ggplot(data.range, aes(x=species_phyla, colour=species_compartm, fill=species_compartm)) +
      geom_bar(stat="count", position=position_dodge(), width= .6) +
      ggtitle("Registered values by species compartment per phyla") + 
      labs(x="", y="Number of registered values") +
      scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 20)) +
      facet_grid(species_type~.) +
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5), legend.position="bottom") +
      theme(legend.title = element_blank())
```
<br>
<br>
<br>

#### Alpha values
Alpha frequency distribution of values registered up to date:

```{r alpha distribution, echo=FALSE, warning=FALSE}
ggplot(data.range, aes(x=alpha))+
      geom_histogram(binwidth=0.1, 
                     breaks = seq(0, 15, by = 0.2), 
                     colour = "black", 
                     fill = "white") +
      geom_rug() +
      geom_density(aes(y=..density..*10), colour="blue") +
      scale_x_continuous(limits=c(0, 15)) +
      scale_y_continuous(limits=c(0, 20)) +
      facet_grid(type_uptake~species_type) +
      ggtitle("Alpha values") + 
      labs(x="Alpha [l g^-1 dw h^-1 µM^-1]", y="Frequency") +
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5))
```
<br>
<br>

*Alpha frequency distribution portrayed as overlapping densities:*

```{r alpha distribution overlapping, echo=FALSE, warning=FALSE}
#Overlapping densities
ggplot(data.range, aes(x=alpha, group=species_type, fill=species_type))+
      geom_density(alpha=0.4) +
      facet_grid(.~type_uptake) +
      ggtitle("Alpha Algae vs Seagrass") + 
      labs(x="Alpha [l g^-1 dw h^-1 µM^-1]", y="Frequency") +
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5)) +
      theme(legend.title = element_blank()) +
      theme(legend.position = c(.9,.88))

```
<br>
<br>
<br>

#### Vmax values
Vmax frequency distribution of values registered  up to date:
<br>

```{r Vmax distribution, echo=FALSE, warning=FALSE}
ggplot(data.range, aes(x=Vmax))+
      geom_histogram(binwidth=0.1, 
                     breaks = seq(0, 15, by = 0.2), 
                     colour = "black", 
                     fill = "white") +
      geom_rug() +
      geom_density(aes(y=..density..*10), colour="blue") +
      scale_x_continuous(limits=c(0, 20)) +
      scale_y_continuous(limits=c(0, 10)) +
      facet_grid(type_uptake~species_type) +
      ggtitle("Vmax values") + 
      labs(x="Vmax [µmol g^-1 dw h^-1]", y="Frequency") +
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5))
```
<br>
<br>
*Vmax frequency distribution portrayed as overlapping densities:*

```{r Vmax distribution overlapping, echo=FALSE, warning=FALSE}
#Overlapping densities
ggplot(data.range, aes(x=Vmax, group=species_type, fill=species_type))+
      geom_density(alpha=0.4) +
      scale_x_continuous(limits=c(0,200)) +
      facet_grid(.~type_uptake) +
      ggtitle("Vmax Algae vs Seagrass") + 
      labs(x="Vmax [µmol g^-1 dw h^-1]", y="Frequency") +
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5)) +
      theme(legend.title = element_blank()) +
      theme(legend.position = c(.9,.88))

```

<br>
<br>

**This concludes this weeks feedback file.**
<br>
<br>
If you have any suggestions or requests for other graphs and information, please let me know.
Best, Simon
